---
title: "Metabolite X"
output: 
  word_document:
    reference_docx: ref.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 300)
```

```{r, echo=FALSE}
metabolite = "metabolite X"
```

# `r metabolite`

```{r load X matrix, echo=FALSE, message=FALSE, warning=FALSE}
library(MetaboMate)
library(hastaLaVista)
library(car)

get.idx=function(range=c(1,5), ppm){
  range=sort(range, decreasing=T);
  which(ppm<=range[1] & ppm>=range[2])}

# X <- "your matrix of data"
X <- bariatricRat.binned.4$binned.data

# x_axis <- "your x-axis"
x_axis <- bariatricRat.binned.4$binned.ppm

# ID <- "a vector with unique ids"
ID <- seq(1, nrow(X))

# group <- country # or the metadata you wish to use for coloring
group <- bariatricRat$metadata$Class

# metadata <- "a data.frame with your metadata"
metadata <- bariatricRat$metadata
```

## Patterns (or combination thereof) used for identification of feature

```{r set pattern and region, echo=FALSE, fig.height=3, fig.width=3.5}
patternID <- 1
identifier = ID
L = length(ID)

# several ranges may be defined and combined to create a pattern
# define here the range that should be used
F <- which(x_axis > 2.42 & x_axis < 2.47)
rangeList <- list(F)
```

From the loadings obtained in block X we can select outliers and use a STOCSY trace to highlight a pattern. We select this region for the `r L` samples in the dataset. A spectra that exhibit a strong feature (the pattern is cleary visible) is chosen as a pattern ID (here: `r patternID`).

```{r ethanol plot pattern, echo=FALSE, fig.height=2.5, fig.width=3.5}
for (i in rangeList) {
  F <- i
  subX <- X[, F]
  
  pattern <- as.numeric(subX[which(identifier == patternID),])
  pattern <- pattern / sum(pattern)
  
  par(mar=c(4,4,2,4))
  title <- paste("pattern for", metabolite)
  plot(x_axis[F], pattern, type = "l", main = title,
       ylab = "arbitrary units",
       xlab = "ppm",
       cex.main = 0.7,
       cex.lab = 0.7,
       cex.axis = 0.7)
}
```
\pagebreak

## Distribution of cross-correlation by class

Each distribution of correlation was obtained using each of the above pattern (or combination thereof) and colour coded by Class. 

```{r compute CCF, echo = FALSE, fig.height=2.5, fig.width=3.5}
# define here what information should be used as a category or a group
colorCode <- metadata$Class

ccList = list()

for (i in seq_along(rangeList)) {
  F <- rangeList[[i]]
  subX <- X[, F]
  
  pattern <- as.numeric(subX[which(identifier == patternID),])
  
  pattern[pattern == 0] <- 1e-04
  dil.F <- 1/apply(subX, 1, function(x) median(x/pattern))
  subX.scaled <- subX * dil.F
  
  res <- apply(subX.scaled, 1, function(x) {
    ccf(pattern, x, lag.max = 10, plot = FALSE, type = c("correlation"))
  })
  
  r <- unlist(lapply(res, function(x) max(x$acf)))
  par(mar=c(4,4,2,4))
  plot(r,
       main = paste0(metabolite, ": cc colored by country (pattern: ", patternID, ")"),
       xlab = "sample index",
       ylab = "cross-correlation",
       cex.main = 0.5,
       cex.lab = 0.5,
       cex.axis = 0.5,
       cex = 0.2,
       col = colorCode)
  ccList[[i]] <- r
}
```

\pagebreak
## Visual inspection of the first 50 samples in the dataset

```{r, echo = FALSE}
# define here the threshold for selection
selectThreshold <- 0.85
# define a threshold for coloring. The samples with cross-correlation below this level (lower confidence) will be displayed with orange color later.
colorThreshold <- 0.9
```

Using the combination of patterns, identification of feature was decided using the threshold of cross correlation (cc) value: high confidence (cc) >`r colorThreshold` was highlighted with green dots in the region; intermediate confidence (cc between `r selectThreshold` to `r colorThreshold`) was highlighted using orange dots and those considered with no features (cc < `r selectThreshold`) using red dots.  

```{r check, echo=FALSE, fig.height=2.5, fig.width=3.5}
# define the spectra region you would like to display for visual inspection (may be larger than the selected pattern)
FL <- get.idx(c(2.35, 2.55), x_axis)
# here you can select a range in case you defined a pattern with multiple ranges
F <- rangeList[[1]]
for (i in 1:50){
  if (r[[i]] > selectThreshold){
    if (r[[i]] < colorThreshold){
      colr = "orange"
    } else {
      colr = "green"
    }
    txt = paste("cc:", round(r[i], 3), "/", metadata$country[i], "/ ID:", identifier[i])
    par(mar=c(4,4,2,4))
    plot(x_axis[FL], as.numeric(X[i, FL]), type = "l", main = txt, col = 1,
         xlab = "ppm", ylab = "arbitrary unit", 
         cex.main = 0.7,
         cex.lab = 0.7,
         cex.axis = 0.7,
         xlim = rev(range(x_axis[FL])),
         ylim = range(X[i, F]))
    points(x_axis[F], as.numeric(X[i, F]), col = colr, cex = 0.2)
  } else {
    txt = paste("cc:", round(r[i], 3), "/", metadata$country[i], "/ ID:", identifier[i])
    plot(x_axis[FL], as.numeric(X[i, FL]), type = "l", main = txt,
         xlab = "ppm", ylab = "arbitrary unit",
         xlim = rev(range(x_axis[FL])),
         ylim = range(X[i, F]),
         cex.main = 0.7,
         cex.lab = 0.7,
         cex.axis = 0.7)
    points(x_axis[F], as.numeric(X[i, F]), col = 2, cex = 0.2)
  }
}
```

\pagebreak

## `r metabolite` statistics

### Percentage of samples with `r metabolite` by Class

```{r stat percent, echo = FALSE}
print(round(table(metadata$Class[which(r > selectThreshold)]) / table(metadata$Class) * 100, 1))
```

### Total number of samples with `r metabolite` by Class

```{r stat, echo = FALSE}
print(table(metadata$Class[which(r > selectThreshold)]))
```

### Total number of samples with `r metabolite`

```{r stat sum, echo = FALSE}
print(sum(table(metadata$Class[which(r > selectThreshold)])))
```




